### 2020-09-14 paper/perfectGraphs.tex
Odd holes algo

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneibhbor in $X$}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

\begin{theorem}

\end{theorem}

To see that the \cref{alg:t3} has a running time of $O(|V|^6)$, let us note that for each triple $v_1, v_2, v_5$ we examine, of which there are $O(|V|^3)$, there are linear many choices of $X$, each taking $O(|V|^2)$ time to process and generating a linear many choices of $v_4$ which take a linear time to process in turn. This gives us the total running time of $O(|V|^6)$.

.

@@

\begin{algorithmic}[1]
	\mProcedure{List-Possible-Near-Cleaners}{$G$}
	\ls $X \gets \emptyset$
	\mForEach{relevant triple $(a, b, c)$}
		\ls $A \gets$ anticomponents of $N(\{a, b\})$
		\mIf{$c$ is $N(\{a, b\})$-complete}
			\ls $r \gets 0$
		\mElse
			\ls $r \gets$ the cardinality of the largest set of $A$ that contains 
			\lsx a nonneibhbor of $c$ 
		\mEndIf
		\ls $Y \gets \emptyset$ 
		\mForEach{$A_i \in A$}
			\mIf{$|V(A_i)| > r$}
				\ls $Y \gets Y \cup A_i$
			\mEndIf
			\mIf{$c \in A_i$}
				\ls $W \gets A_i$
			\mEndIf
		\mEndFor
		\ls $Z \gets$ the set of all $Y \cup W$-complete vertices
		\ls $X \gets X \cup (Y \cup Z)$ 
	\mEndFor
	\ls $N \gets \emptyset$
	\mForEach{edge $uv \in E(G)$}
		\ls $N \gets N \cup N(\{u, v\})$
	\mEndFor
	\ls $R \gets \emptyset$
	\mForEach{$X_i \in X$}
		\mForEach{$N_j \in N$}
			\ls $R \gets R \cup (X_i \cup N_j)$
		\mEndFor
	\mEndFor
	\ls \RETURN $R$
	\mEndProcedure
\end{algorithmic}

@@

\begin{theorem}[9.1 of \cite{MC05}]
	\label{thm:91}
	Let $C$ be a shortest odd hole in $G$, with length at least 7. Then there is a relevant triple $(a, b, c)$ of vertices such that
	\begin{itemize}
		\item the set of all $C$-major vertices not in $X(a, b, c)$ is anticonnected
		\item $X(a, b, c) \cap V(C)$ is a subset of the vertex set of some 3-vertex path of $C$.
	\end{itemize}
\end{theorem}

Let us suppose that $C$ is an amenable hole in $G$. By \cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c)$ is a near-cleaner for $C$. Therefore the output of the \cref{alg:listNearCleaners} is correct.

@@

	\noindent Output: Determines one of the following:
	\begin{itemize}
		\item $G$ has an odd hole
		\item There is no shortest odd hole $C$ such that $X$ is a near-cleaner for $C$.
	\end{itemize}
\end{alg}
\begin{algtext2}
	For every pair $x, y \in V(G)$ of distinct vertices find shortest path $R(x, y)$ between $x, y$ with no internal vertex in $X$. If there is one, let $r(x, y)$ be its length, if not, let $r(x, y) = \infty$.

	For all $y_1 \in V(G)\setminus X$ and all 3-vertex paths $x_1-x_2-x_3$ of $G\setminus y_1$ we check the following:
	\begin{itemize}
		\item $R(x_1, y_1), R(x_2, y_2)$ both exist -- define $y_2$ as the neighbor of $y_1$ in $R(x_2, y_1)$.
		\item $r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2) = min(r(x_3, y_1), r(x_3, y_2))$ 
	\end{itemize}

	If there is such a choice of $x_1$, $x_2$, $x_3$, $y_1$ then we output that there is an odd hole. If not, we report that there is no shortest odd hole $C$ such that $X$ is a near-cleaner for $C$.
\end{algtext2}

\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/nearCleanerFound.tex}
	\caption{An odd hole is found}%
	\vspace{-0.5cm}
\end{wrapfigure}

@@

\begin{proof}
	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

@@

	Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq n (= r(x_1, y_2))$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = n$ we get that $y_2$ does not belong to $R(x_1, y_1)$.

@@

>>>>

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneibhbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

\begin{theorem}
	\Cref{alg:t3} runs in $O(|V|^6)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triples $(v_1, v_2, v_5)$ we enumerate in line \ref{line:t3start}. For each, there are $O(|V|)$ $X$s, so lines \ref{line:t3n4start}-\ref{line:t3n4end} are executed $O(|V|^4)$ times.

	Each execution of lines \ref{line:t3n4start}-\ref{line:t3mid} takes $O(|V|^2)$ time, and there are $O(|V|)$ $v_4$s to enumerate. Line \ref{line:t3v4if} takes $O(|V|)$ time and then there are $O(|V|)$ $v_3$s to enumerate. Each is checked in $O(1)$ time. If we find the correct one, we calculate a path $P$, but return right away, so we calculate $P$ at most once in whole algorithm.
\end{proof}

@@

\begin{algorithmic}[1]
	\mProcedure{List-Possible-Near-Cleaners}{$G$}
	\ls $\mathcal{X} \gets \emptyset$
	\mForEach{relevant triple $(a, b, c)$}
		\ls $A \gets$ anticomponents of $N(\{a, b\})$
		\mIf{$c$ is $N(\{a, b\})$-complete}
			\ls $r \gets 0$
		\mElse
			\ls $r \gets$ the cardinality of the largest set of $A$ that contains 
			\lsx a nonneibhbor of $c$ 
		\mEndIf
		\ls $Y \gets \emptyset$ 
		\mForEach{$A_i \in A$}
			\mIf{$|V(A_i)| > r$}
				\ls $Y \gets Y \cup A_i$
			\mEndIf
			\mIf{$c \in A_i$}
				\ls $W \gets A_i$
			\mEndIf
		\mEndFor
		\ls $Z \gets$ the set of all $Y \cup W$-complete vertices
		\ls $X(a, b, c) \gets Y \cup Z)$ \Comment{we use $X(a, b, c)$ in the proof of correctness}
		\ls $\mathcal{X} \gets \mathcal{X} \cup X(a, b, c)$
	\mEndFor
	\ls $\mathcal{N} \gets \emptyset$
	\mForEach{edge $uv \in E(G)$}
		\ls $\mathcal{N} \gets \mathcal{N} \cup N(\{u, v\})$
	\mEndFor
	\ls $\mathcal{R} \gets \emptyset$
	\mForEach{$X_i \in X$}
		\mForEach{$N_j \in N$}
			\ls $\mathcal{R} \gets \mathcal{R} \cup (X_i \cup N_j)$
		\mEndFor
	\mEndFor
	\ls \RETURN $\mathcal{R}$
	\mEndProcedure
\end{algorithmic}

@@

\begin{theorem}[9.1 of \cite{MC05}]
	\label{thm:91}
	Let $C$ be a shortest odd hole in $G$, with length at least seven. Then there is a relevant triple $(a, b, c)$ of vertices such that
	\begin{itemize}
		\item the set of all $C$-major vertices not in $X(a, b, c)$ is anticonnected
		\item $X(a, b, c) \cap V(C)$ is a subset of the vertex set of some 3-vertex path of $C$.
	\end{itemize}
\end{theorem}

\begin{theorem}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \cref{alg:listNearCleaners} is correct.
\end{proof}

@@

	\noindent Output: Determines one of the following:
	\begin{itemize}
		\item $G$ has an odd hole
		\item There is no shortest odd hole $C$ such that $X$ is a near-cleaner for $C$.
	\end{itemize}
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2) =$\lsx$min(r(x_3, y_1), r(x_3, y_2))$}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf
				\mEndFor
			\mEndFor
		\mEndFor
		\RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

@@

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole.
\end{theorem}
\begin{proof}
	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.3\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		% \vspace{-0.2cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1-\ldots -p_n$, and let $R(x_2, y_1) = q_1-\ldots -q_{n+1}$, where $p_1 = x_1$, $p_n = q_{n+1} = y_1$, $q_1 = x_2$ and $q_n = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{n-1}, q_2, \ldots, q_n$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

@@

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq n (= r(x_1, y_2))$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = n$ we get that $y_2$ does not belong to $R(x_1, y_1)$.\par}%

@@

\begin{wrapfigure}{r}{0.35\textwidth}

\end{wrapfigure}
